Predr_Cors

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Aims.Rmd
wip

(i)	criteria for assessing ethical CHEM implementations;

@@

(iii)	use of the software framework to develop an open-source CHEM in youth mental health, with an initial focus on outcome valuation.

>>>>

(i)	proposed criteria for assessing ethical CHEM implementations;

@@

(iii)	how the software framework is being used to develop an open-source CHEM in youth mental health, with an initial focus on valuing health outcomes.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Discussion.Rmd
wip

<!-- Talk about DARTH (https://doi.org/10.1007/s40273-019-00837-x )/ BCEA -->
Ethical practice is a core expectation of health researchers and computational methods underpin most quantitive research, yet an understanding of what constitutes ethical computational modelling practice in health is underdeveloped [@10.3389/fpubh.2017.00068]. The modeller responsibilities, enabling model attributes and model implementation assessment criteria that we propose can help address this gap.

The ethical responsibilities and enabling model attributes we describe have both commonalities and distinctive features compared to a previous ethical framework for computational modelling in public health [@10.3389/fpubh.2017.00068]. The authors of that framework propose 13 questions to evaluate ethical risk across the four criteria of independence, transparency, benficience and justice. Their descriptions of the four criteria at least partially map to either our proposed modeler responsibilities (“justice” to “social acceptability”, “independence” to “adequacy for purpose” and “beneficience” to “beneficial impact”) or enabling model attributes (“transparency”). However, while our six assessment criteria are specific to three attributes (TRU) of the computational implementation of the model, the prior ethical framework includes questions relevant to the the conceptual and mathematical models and the potential impacts of model use. Examples of these more general evaluation questions include (for the justice criterion) “is any lack of knowledge about important parameters attributable to uncertainty or variability?” and (for the beneficience criterion) “if a policy is based on the model evidence, is it more likely to be effective and beneficial than a decision made in the absence of the model?”. The less numerous and more focused assessment criteria we propose may potentially be more tractable to implement in reviews of models authored by third parties and as the basis for designing software frameworks to support ethical computational model implementation.

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reusable [@Feenstra2022; @Emerson2019] and updatable [@Sampson2017; @10.3389/fpubh.2022.899874]]. <!-- DELETION --> Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work. Currently, it takes “an extraordinary amount of idealism” to dedicate the substantial time and resources required to author, test, document and maintain even fragile prototype research software that could instead be used to write scientific manuscripts [@anzt2020environment].

@@

Our prototype framework has a number of features that subsequent work to develop ethical software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R or python, can keep framework scope relatively narrow (making it more tractable to develop, maintain and learn) while readily leveraging and coherently integrate other modelling tools written in that language (e.g. the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts (a limitation of object oriented approaches famously described as: “you wanted a banana but what you got was a gorilla holding the banana” [@seibel2009joe]). Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritizing the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code naming conventions and file organisation which in turn enables automated generation of simple documentation at every code update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customized documentation less frequently.

>>>>

<!-- Talk about DARTH (https://doi.org/10.1007/s40273-019-00837-x )/ BCEA -->
Ethical practice is a core expectation of health researchers and computational methods underpin most quantitative research, yet an understanding of what constitutes ethical computational modelling practice in health is underdeveloped [@10.3389/fpubh.2017.00068]. The modeller responsibilities, enabling model attributes and model implementation assessment criteria that we propose can help address this gap.

The ethical responsibilities and enabling model attributes we describe have both commonalities and distinctive features compared to a previous ethical framework for computational modelling in public health [@10.3389/fpubh.2017.00068]. The authors of that framework propose 13 questions to evaluate ethical risk across the four criteria of independence, transparency, beneficence and justice. Their descriptions of the four criteria at least partially map to either our proposed modeler responsibilities (“justice” to “social acceptability”, “independence” to “adequacy for purpose” and “beneficence” to “beneficial impact”) or enabling model attributes (“transparency”). However, while our six assessment criteria are specific to three attributes (TRU) of the computational implementation of the model, the prior ethical framework includes questions relevant to the conceptual and mathematical models and the potential impacts of model use. Examples of these more general evaluation questions include (for the justice criterion) “is any lack of knowledge about important parameters attributable to uncertainty or variability?” and (for the beneficence criterion) “if a policy is based on the model evidence, is it more likely to be effective and beneficial than a decision made in the absence of the model?”. The less numerous and more focused assessment criteria we propose may potentially be more tractable to implement in reviews of models authored by third parties and as the basis for designing software frameworks to support ethical computational model implementation.

Currently, many if not most existing CHEMs are insufficiently transparent [@Jalali2021; @McManus2019; @Bermejo2017; @Ghabri2019], reusable [@Feenstra2022; @Emerson2019] and updatable [@Sampson2017; @10.3389/fpubh.2022.899874]]. <!-- DELETION --> Existing incentive structures for health economists generally do not promote facilitating peers to reuse their work. Currently, it can take “an extraordinary amount of idealism” to commit to authoring and maintaining research software [@anzt2020environment].

@@

Our prototype framework has a number of features that subsequent work to develop ethical software frameworks may find useful to incorporate. Firstly, developing a software framework to work within an existing and widely used open source programming language such as R or python, can keep framework scope relatively narrow (making it more tractable to develop, maintain and learn) while readily leveraging and coherently integrate other modelling tools written in that language (e.g. the dependency libraries we list in Table \@ref(tab:cpkgs)). Secondly, implementation that combines both object oriented and functional programming paradigms can avail of the modular and syntactical simplicity benefits of the former, while limiting needless bundling of code artefacts. Thirdly, a sensible trade-off needs to be found between transparent code implementation (which requires clear and sufficiently detailed documentation) and Agile Software Development (for which a foundational principle is prioritizing the development of working code over writing documentation [@beck2001manifesto]). Our software framework makes this trade off by enforcing the use of consistent code naming conventions and file organisation which in turn enables automated generation of simple documentation at every code update. All model data-structures and algorithms are therefore always documented (at least minimally, with machine authored content), meaning model developers have a requirement to write customized documentation less frequently.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Imp_Example.Rmd
about to start on updating example
wip

##	Economic topics
Currently, we are using the ready4 software framework to develop, apply and share youth mental health computational models in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

@@

-	supply of health services (our projects: a model of primary mental health care services).

Once these projects are completed, our aim is to flexibly combine these models as modules of a multi-purpose model (also called “ready4”) to answer questions in two additional Wagstaff and Culyer domains:

-	efficiency and equity (our goal: assess the distributional impacts and identify the optimal targeting of care provision); and

@@

## Case study: Utility mapping
Our initial application of the ready4 software framework was to undertake a previously described study [@Hamilton2021.07.07.21260129] to develop utility mapping models for use in samples of young people presenting to primary mental health services. The ready4 software framework was used to develop CHEM modules, supply those modules with data and implement modelling analyses, creating the following artefacts:

@@

 - data collections of synthetic populations for testing model modules [@DVN/HJXYKQ_2021] and study input and results data [@DVN/DKDIB0_2021];

 - programs for replicating all steps from data ingest to manuscript reporting [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821];

@@

We created a checklist (Table \@ref(tab:checktb)) that we used to subjectively assess these study outputs against TRU criteria. For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe the outputs from our utility mapping study may be assessable as having satisfactorily met four of the six criteria (T1, T2, R2 and U1) and to have partially met two criteria (R1 and U2). The main shortcomings that we identified when applying the assessment criteria were the need for additional development before the CHEM modules would be sufficiently generalizable for valid application in datasets that measure health utility with different instruments and a general lack of unit testing.

>>>>

<!-- Add intro text -->
We plan on using the ready4 software framework to implement multiple transparent, reusable and updatable CHEMs in youth mental health (see Figure \@ref(fig:fig1)).  Some initial outputs from this work are publicly available.

```{r, child = child_figs_ls$fig1, echo=FALSE}
```

##	Economic topics
Currently, we are using the ready4 software framework to develop, apply and share youth mental health CHEMs in four of the twelve domains of health economics identified by Wagstaff and Culyer [@wagstaff2012four]:

@@

-	supply of health services (our projects: a model of primary mental health care services for young people).

Potential future directions are to supplement this work with CHEMs in two additional Wagstaff and Culyer domains of public health (to model the impact of selected fiscal policy and regulation options on young people's mental health) and human resources (to model the supply and behaviours of the youth mental health workforce). Our ultimate aim is to flexibly combining all our CHEMs in analyses that help answer questions in two additional Wagstaff and Culyer domains:

-	efficiency and equity (our goal: assess the distributional impacts and identify the optimal targeting of youth mental health interventions); and

@@

Although principally interested in using our CHEMs to answer policy questions relating to the mental health of young people in Australia, we want to facilitate CHEM transferability to other jurisdictions. Our CHEMs are therefore being derived from and applied to real data (which can be assumption, empirical or simulated, so long as it is appropriate for use in analysis intended to inform decision-making) from Australia. Additionally, we plan for all our CHEMs to be distributed with toy (fake) data to help demonstrate the potential use of these CHEMS in other decision contexts.

## Case study: health and its value
We have previously described [@Hamilton2021.07.07.21260129] a study to develop utility mapping models for use in samples of young people presenting to primary mental health services. The ready4 software framework was used in that study to develop CHEM modules, supply those modules with data and implement modelling analyses, creating the following artefacts:

@@

 - collections of real data (study input and results [@DVN/DKDIB0_2021]) and fake data (synthetic populations for testing model modules [@DVN/HJXYKQ_2021]);

 - programs for replicating all steps from data ingest to manuscript reporting [@hamilton_matthew_2022_6129906], applying utility mapping models to new data [@hamilton_matthew_2022_6416330] and generating a synthetic representation of the study dataset [@hamilton_matthew_p_2022_6321821]; and

@@

We created a checklist (Table \@ref(tab:checktb)) that we used to subjectively assess these study outputs against TRU criteria. For each criterion, we provided a global assessment of whether it was met using the responses “yes”, “no” or “partial”.  We believe the outputs from our utility mapping study may be assessable as having satisfactorily met five of the six criteria (T1, T2, R1, R2 and U1) and to have partially met one criterion (U2). The main shortcomings that we identified when applying the assessment criteria was that we have yet to adequately implement unit testing of the R libraries we authored as part of this study.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Imp_Guidelines.Rmd
wip

We considered prior literature on modelling practice, our own professional experience and the needs of modelling projects we are undertaking to identify: (i) some core ethical responsibilities of CHEM developers; (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.

## Ethical responsibilities of CHEM developers 
We suggest that three core ethical responsibilities of health economic modellers are to make **socially acceptable** value judgments when planning a CHEM project, to ensure that implementation of a CHEM is **fit for purpose** and to promote **socially beneficial** use of the CHEM once it is released. There is significant scope for health economists to improve fulfilment of each of these responsibilities.

@@

A CHEM that is **reusable** also signals ethical modelling practice. Making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, yet is not sufficient to promote model re-use.  Key concepts in health economic model reuse include generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are easier to selectively modify.  Object-oriented programming approaches incorporate a concept called inheritance [@8717448] that can allow duplicate models to be created and selectively modified. Modular models are constructed from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. To grant permissions to others to use, test and adapt models and their components, health economists can avail of two broad categories of open source licensing options. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

@@

##	Assessment criteria
For each attribute that suggests ethical CHEM implementations, we identified assessment criteria that can be met using existing technical solutions and which therefore have the potential to be facilitated by software frameworks.

>>>>

We considered prior literature on modelling practice and our own professional experience to identify: (i) some core ethical responsibilities of CHEM developers; (ii) attributes of CHEMs that can suggest fulfillment of these responsibilities; and (iii) criteria against which these attributes can be assessed.

## Ethical responsibilities of CHEM developers 
For simplicity, we sought to identify one overarching ethical responsibility for modellers at each distinct phase of a CHEM project. As CHEMs are implemented as software, software development lifecycle (SDLC) models may provide an appropriate basis for identifying the main phases of a CHEM project. There are multiple SDLC models, each taking a different approach to the number, naming, definition, sequencing and iteration of project phases [@ruparelia2010software]. However, components of most SDLC models can broadly be mapped to the concepts of *planning* (in which software requirements and design specification are agreed), *development* (in which software is authored and tested) and *deployment* (in which software is made available as production ready tools for end-users). We contend that ethical responsibilities of computational modellers identified in prior literature [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational] can be summarised under the following three overarching goals (one for each phase):

- **make socially acceptable value judgments** (when planning CHEM projects);

- **ensure model fitness for purpose** (when developing CHEMs); and

- **promote socially beneficial use of models** (once CHEMs have been deployed).

There is significant scope for health economists to improve fulfilment of each of these responsibilities.

@@

A CHEM that is **reusable** also signals ethical modelling practice. Making a CHEM’s code, data and documentation publicly available is increasingly considered good practice, yet is not sufficient to promote model re-use.  Key concepts in health economic model reuse include generalisability (application without adaptation) and transferability (selective reuse and/or modification of model components) [@RN39]. Writing model algorithms as collections of functions (short, self-contained and reusable software routines that each perform a discrete task) is good scientific computing practice [@Wilson_2017] and promotes selective reuse. Computational implementations that store model code and data in distinct files and locations (as opposed to embedding data such as parameter values into source code) are easier to selectively modify. Modular implementations construct models from multiple independently reusable and replaceable sub-models (modules) [@pan2021modular]. The programming concept inheritance [@8717448], deployed in object-oriented programming approaches, can facilitate the duplication and selective modification of models. Granting permissions to others to use, test and adapt models and their components, can be facilitated by two broad categories of open source licenses. Some guidance strongly recommends the use of permissive licensing [@Wilson_2017] that provides users with great flexibility as to the purposes (including commercial) for which content can be re-used. An alternative approach is to use copyleft licenses [@copyleft2022] that can require content users to distribute any derivative works they create under similar open-source arrangements. 
<!-- Compared to using commercial modelling software, authoring CHEMs in an open-source language like R [@RCORE2022] makes it easier to -->
<!-- Using open-source development platforms and licenses can aid both generalizability and transferability. -->
<!-- The choices that CHEM developers make about model implementation and licensing will also shape who can use a model and for what purposes.  -->
<!-- are relatively easier to selectively modify. -->

@@

##	Assessment criteria
For each CHEM attribute that suggests ethical CHEM modelling practice, we identified assessment criteria that have the potential to be met with the assistance of software frameworks.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Imp_Toolkit.Rmd
wip

## R libraries 
A library in the R language will typically depend on multiple other R libraries, each of which may have different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g. the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One framework library provides a **foundation for modular CHEM implementations**. In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. To achieve this, the foundation framework library defines a template module (using R’s S4 class system), which can be used to create other modules with a common set of inherited properties. One of these inherited properties is a novel syntax of 15 core commands that enable module algorithms to be consistently named. The foundation library also contains tools for retrieving web-based information on model modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

Three framework libraries are designed to help streamline and standardise workflows for **authoring CHEM modules** from the template module. The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Authoring with each of these paradigms is facilitated by a dedicated CHEM module authoring library. The first module authoring library contains tools for writing functions in a consistent house style and then using the standardised naming conventions of that house style to automatically generate basic documentation for each function. The second module authoring library contains tools to help streamline and standardise the authoring of novel CHEM modules with distinctive properties and for associating novel algorithms with these modules using the framework's syntax. The second module authoring library also supports the automatic generation of basic documentation for each new CHEM module. The third module authoring library provides tools for disseminating themed bundles of CHEM modules as R libraries that are:

@@

We created a GitHub organisation (a collection of code repositories) where all code (libraries, programs and sub-routines) that we author with the framework is stored and version controlled. We configured individual repositories in our GitHub organisation that are used for code library projects to use GitHub actions to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository used for code library development.

>>>>

## R libraries 
A library in the R language will typically depend on multiple other R libraries, all of which potentially having different authors. As the number of third-party dependencies of an R library grows so does the fragility of that library (e.g. the library may cease to work as intended due to changes in one of its dependency libraries). To reduce the fragility of our framework we implemented it as multiple R libraries rather than one R library. In total we authored six novel R libraries to implement the ready4 framework, all of which have distinct purposes and dependencies (Table \@ref(tab:cpkgs)).

One framework library provides a **foundation for modular CHEM implementations**. In modular model implementations, modules need to be able to share inputs and outputs with each other and to be run as independent models [@barros2023empowering]. To achieve this, the foundation framework library defines a template CHEM module (using R’s S4 class system), which can be used to create other CHEM modules with a common set of inherited properties. One of these inherited properties is a novel syntax of 15 core commands that enable CHEM module algorithms to be consistently named. The foundation library also contains tools for retrieving web-based information on CHEM modules, datasets and analysis programs authored with the framework and for partially automating updates to a project documentation website.

Three framework libraries are designed to help streamline and standardise workflows for **authoring CHEM modules** from the template module. The R language supports functional and object-oriented programming paradigms [@10.1214/13-STS452]. Authoring with each of these paradigms is facilitated by a dedicated CHEM module authoring library. One module authoring library contains tools for writing functions in a consistent house style and then using the standardised naming conventions of that house style to automatically generate basic documentation for each function. A second module authoring library contains tools to help streamline and standardise the authoring and documenting of novel CHEM modules. A third module authoring library provides tools for disseminating themed bundles of CHEM modules as R libraries that are:

@@

We created a GitHub organisation (a collection of code repositories) where code (libraries, programs and sub-routines) that we author for and with the framework is stored and version controlled. We configured individual repositories in our GitHub organisation to use GitHub actions to implement continuous integration. By default, code libraries authored with our framework will use continuous integration to assess compliance with policies specified by the Comprehensive R Archive Network (CRAN) [@CRAN2022]. To track our code coverage, we linked our GitHub organisation to an account we established at codecov [@codecov_2022]. To facilitate the creation and hosting of documentation websites, we enabled GitHub Pages in each repository used for code library development.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Intro_Short.Rmd
wip

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four].  In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and computer code (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., through execution of a computer program that applies the computational model to selected input data). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. Independent assessment of the validity of this assumption depends in part on how a computational model has been authored and shared.

@@

Computational health economic models (CHEMs) can be implemented using specialized commercial software or authored as software development projects in an open-source programming language such as Python [@python2009] or R [@RCORE2022]. Advantages of commercial modelling tools include simplicity and ease of use, but an open-source software development approach may facilitate computational models that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what type of implementation to choose for a specific CHEM project needs to balance both desired model features and the resource and time constraints within which the project needs to be delivered (see Table \@ref(tab:proscons)).

```{r, child = child_docs_ls$one, echo=FALSE}
```
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
For open-source CHEM projects, a major consideration in is the selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources, e.g., code and data files, that can be modified to suit specific needs. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

@@

<!-- We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.   -->
<!-- When making choices about how to implement this health economic model computationally, we wished to facilitate ethical development and use. -->
The ethical responsibilities of computational modellers are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational]. Guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce. A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. However, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019], and novel software frameworks may be necessary to help health economists fulfill a broader set of ethical responsibilities.  Ideally, software frameworks for ethical CHEMs would provide tools that address known barriers to ethical computational modelling practice. The main barriers to open-source CHEMs identified in a 2020 survey of health economists concern updating models, legal considerations, transferring model data, the platforms or languages in which models are developed, the level of public access required, lack of interest from decision-makers, model generalisability and confidentiality and security considerations [@Pouwels2022].

>>>>

Health economics is a discipline concerned with problems that arise due to scarce resources, such as how to value health and healthcare, allocate healthcare budgets and configure health services [@wagstaff2012four]. In seeking to solve these problems, health economists typically use models which are simplified and selective representations of systems that are believed to influence human health. A health economic model should be capable of being described using words and figures (a conceptual model), equations (a mathematical model) and computer code (a computational model).  Health economic scientific manuscripts typically describe a model using its conceptual and mathematical representations but report results that have been generated by its computational representation (i.e., through execution of a computer program that applies the computational model to selected input data). The conceptual, mathematical and computational representations of a model are assumed to be isomorphic. However, the ease of independently assessing the validity of this assumption depends in part on how a health economic model has been implemented computationally.

@@

Computational health economic models (CHEMs) can be implemented using specialised commercial modelling software or authored as software development projects in an open-source programming language such as Python [@python2009] or R [@RCORE2022]. Advantages of commercial modelling tools include simplicity and ease of use, but an open-source software development approach may facilitate computational models that are more transparent, reusable and updatable [@incerti2019r; @Pouwels2022]. The decision about what implementation approach to adopt needs to balance both modelling team value judgments about priority model features and the resource and time constraints within which a CHEM project needs to be delivered (see Table \@ref(tab:proscons)).

```{r, child = child_docs_ls$one, echo=FALSE}
```
<!-- on considerations such as model comprehensiveness, generalisability, reproducibility and transferability -->
For open-source software projects, a major consideration is the selection of an appropriate software framework. A software framework is a shared common technology used by developers to collaboratively author software and which is not typically visible to software end-users [@myllarniemi2018development]. A software framework provides a foundation for developing multiple software applications with shared resources (e.g., code and data files), that can be modified to suit specific needs. Software frameworks have widely been developed and implemented in data science, for example, PyTorch [@PyTorch2019] for machine learning using Python.

@@

<!-- We are developing a model to explore multiple economic questions relating to the mental health of young people aged 12 to 25.   -->
<!-- When making choices about how to implement this health economic model computationally, we wished to facilitate ethical development and use. -->
The ethical responsibilities of computational modellers seeking to inform policy decisions are often poorly understood and inadequately fulfilled [@10.3389/fpubh.2017.00068; @pliakos2021ethics; @thompson2022escape; @calder2018computational]. Within the health economics discipline, open-source CHEMs remain rare [@Feenstra2022; @Emerson2019]. The main barriers to open-source CHEMs identified in a 2020 survey of health economists concern challenges relating to updating models, legal considerations, transferring model data, the platforms or languages in which models are developed, the level of public access required, lack of interest from decision-makers, model generalisability and confidentiality and security considerations [@Pouwels2022].

There is a potential role for software frameworks to help health economic modellers overcome some of these barriers. A high-level software framework for implementing open-source CHEMs in R has previously been developed with the primary aim of improving model transparency [@alarid2019need]. Ideally, software frameworks would facilitate fulfillment of a broader set of ethical responsibilities. However, guidance on developing, selecting and using software frameworks for ethical CHEM implementations is scarce.

### 2023-10-08 Markdown/Manuscript/Child_RMDs/MS_Tables_2.Rmd
wip

```{r timelygls, tab.cap='Framework standards', tab.id = 'timelygls', echo = F, results = 'asis'}

>>>>

```{r timelygls, tab.cap='How transparent, reusable and updatable (TRU) computational health economic models (CHEMs) suggest ethical modelling practice', tab.id = 'timelygls', echo = F, results = 'asis'}
